---
title: "MAT5314 Project 1: Data Visualization"
author: 
  - Teng Li(7373086)
  - Shiya Gao(300381032) 
  - Chuhan Yue(300376046)
  - Yang Lyu(8701121)
output: pdf_document
header-includes: 
  - \renewcommand{\and}{\\}
---

# Introduction
A data set of the 2016 US election polls was given. In this project we aim to understand the data structure by creating various visualizations.  

# Method
We use various R packages to present the data set and to plot the graphs.
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(tidyverse)
library(ggplot2)
```
# Result
We first take a look at the raw data set:
```{r}
ElectionPoll<-read.csv("polls_us_election_2016.csv", header = TRUE)
head(ElectionPoll)
tail(ElectionPoll)
```

As we can see, there are a few variables with missing values:
```{r}
summary(ElectionPoll)
```

We first explore the column name of raw data.
```{r}
View(ElectionPoll)
# Explore column name of raw data
column_name <- colnames(ElectionPoll)
print(column_name)
```

The following is the values in state:
```{r}
# Explore values in state
print(unique(ElectionPoll$state))
```

Count the missings in state
```{r}
# Count the missings in state
missing_state <- sum(is.na(ElectionPoll$state))
print(missing_state)
```

The following is the values in startdate:
```{r}
# Explore unique in startdate
sort(unique(ElectionPoll$startdate))
```

Count the missings in startdate
```{r}
# Count the missings in startdate
missing_startdate <- sum(is.na(ElectionPoll$startdate))
print(missing_startdate)
```

The following is the values in enddate:
```{r}
# Explore unique in enddate
sort(unique(ElectionPoll$enddate))
```

Count the missings in enddate
```{r}
# Count the missings in enddate
missing_enddate <- sum(is.na(ElectionPoll$enddate))
print(missing_enddate)
```

The following is the values in pollster:
```{r}
# Explore values in pollster
print(unique(ElectionPoll$pollster))
```

Count the missings in pollster
```{r}
# Count the missings in pollster
missing_pollster <- sum(is.na(ElectionPoll$pollster))
print(missing_pollster)
```

The following is the values in grade:
```{r}
# Explore values in grade
print(unique(ElectionPoll$grade))
```

Count the missings in grade
```{r}
# Count the missings in grade
missing_grade <- sum(is.na(ElectionPoll$grade))
print(missing_grade)
```

The following is the values in population:
```{r}
# Explore values in population
print(unique(ElectionPoll$population))
# Count the number
table(ElectionPoll$population)
```

Count the missings in population
```{r}
# Count the missings in population
missing_population <- sum(is.na(ElectionPoll$population))
print(missing_population)
```

The following is the number of unique and size in rawpoll_clinton:
```{r}
# Explore values in rawpoll_clinton
length(unique(ElectionPoll$rawpoll_clinton))
# Size of rawpoll_clinton
length(ElectionPoll$rawpoll_clinton)
```

Count the missings in rawpoll_clinton
```{r}
# Count the missings in rawpoll_clinton
missing_rawpoll_clinton <- sum(is.na(ElectionPoll$rawpoll_clinton))
print(missing_rawpoll_clinton)
```

The following is the number of unique and size in rawpoll_trump:
```{r}
# Explore values in rawpoll_trump
length(unique(ElectionPoll$rawpoll_trump))
# Size of rawpoll_trump
length(ElectionPoll$rawpoll_trump)
```

Count the missings in rawpoll_trump
```{r}
# Count the missings in rawpoll_trump
missing_rawpoll_trump <- sum(is.na(ElectionPoll$rawpoll_trump))
print(missing_rawpoll_trump)
```

The following is the number of unique and size in rawpoll_johnson:
```{r}
# Explore values in rawpoll_johnson
length(unique(ElectionPoll$rawpoll_johnson))
# Size of rawpoll_johnson
length(ElectionPoll$rawpoll_johnson)
```

Count the missings in rawpoll_johnson
```{r}
# Count the missings in rawpoll_johnson
missing_rawpoll_johnson <- sum(is.na(ElectionPoll$rawpoll_johnson))
print(missing_rawpoll_johnson)
```

The following is the number of unique and size in rawpoll_mcmullin:
```{r}
# Explore values in rawpoll_mcmullin
length(unique(ElectionPoll$rawpoll_mcmullin))
# Size of rawpoll_mcmullin
length(ElectionPoll$rawpoll_mcmullin)
```

Count the missings in rawpoll_mcmullin
```{r}
# Count the missings in rawpoll_mcmullin
missing_rawpoll_mcmullin <- sum(is.na(ElectionPoll$rawpoll_mcmullin))
print(missing_rawpoll_mcmullin)
```

The following is the number of unique and size in adjpoll_clinton:
```{r}
# Explore values in adjpoll_clinton
length(unique(ElectionPoll$adjpoll_clinton))
# Size of adjpoll_clinton
length(ElectionPoll$adjpoll_clinton)
```

Count the missings in adjpoll_clinton
```{r}
# Count the missings in adjpoll_clinton
missing_adjpoll_clinton <- sum(is.na(ElectionPoll$adjpoll_clinton))
print(missing_adjpoll_clinton)
```

The following is the number of unique and size in adjpoll_trump:
```{r}
# Explore values in adjpoll_trump
length(unique(ElectionPoll$adjpoll_trump))
# Size of adjpoll_trump
length(ElectionPoll$adjpoll_trump)
```

Count the missings in adjpoll_trump
```{r}
# Count the missings in adjpoll_trump
missing_adjpoll_trump <- sum(is.na(ElectionPoll$adjpoll_trump))
print(missing_adjpoll_trump)
```

The following is the number of unique and size in adjpoll_Johnson:
```{r}
# Explore values in adjpoll_johnson
length(unique(ElectionPoll$adjpoll_johnson))
# Size of adjpoll_johnson
length(ElectionPoll$adjpoll_johnson)
```

Count the missings in adjpoll_Johnson
```{r}
# Count the missings in adjpoll_Johnson
missing_adjpoll_johnson <- sum(is.na(ElectionPoll$adjpoll_johnson))
print(missing_adjpoll_johnson)
```

The following is the number of unique and size in adjpoll_mcmullin:
```{r}
# Explore values in adjpoll_mcmullin
length(unique(ElectionPoll$adjpoll_mcmullin))
# Size of adjpoll_mcmullin
length(ElectionPoll$adjpoll_mcmullin)
```

Count the missings in adjpoll_mcmullin
```{r}
# Count the missings in adjpoll_mcmullin
missing_adjpoll_mcmullin <- sum(is.na(ElectionPoll$rawpoll_mcmullin))
print(missing_adjpoll_mcmullin)
```

Data dictionary:  
* state: The name of the state where the election is held, “U.S.” means national polls  
* startdate: Start data of poll  
* enddate: End data of poll  
* pollster: Organization name that conducts or analyzes opinion polls  
* grade: Grade assigned by Fivethirtyeight (sometimes as 538, an American website that focuses on opinion poll) to pollster  
      >"A letter grade from A+ to F that reflects the accuracy of a polling organization's polls and its predictive plus-minus score (a projection of how well we think a pollster's polls will do in the future). Pollsters with a small number of polls get a provisional rating (for example, A/B) rather than a precise letter grade“ (Cite from https://projects.fivethirtyeight.com/pollster-ratings/).  
* samplesize: Sample size of polls for each pollster  
* population: Type of population being polled  
            *1.LV (likely Voters): population whose most likely to participate in 2016 election.  
            2.RV (Registered Voters): RV includes all eligible and registered voters, whether or not they end up voting.  
            3.A (Adults): Refers to all persons of legal voting age.  
            4.V (Voters): Persons who has voted in an election.  
* rawpoll_clinton: Percentage for Hillary Clinton  
* rawpoll_trump: Percentage for Donald Trump  
* rawpoll_johnson: Percentage for Gary Johnson
* rawpoll_mcmullin: Percentage for Evan Mcmullin
* adjpoll_clinton: Fivethirtyeight adjusted percentage for Hillary Clinton  
* adjpoll_trump: Fivethirtyeight adjusted percentage for Donald Trump  
* adjpoll_johnson: Fivethirtyeight adjusted percentage for Gary Johnson  
* adjpoll_mcmullin: Fivethirtyeight adjusted percentage for Evan Mcmullin  

```{r}
ggplot(data = ElectionPoll, mapping = aes(x = grade, y = rawpoll_trump)) +
    geom_boxplot()
```
```{r}

```

# Discussion

# Conclusion
