---
title: "MAT5314 Project 1: Data Visualization"
author: 
  - Teng Li(7373086)
  - Shiya Gao(300381032) 
  - Chuhan Yue(300376046)
  - Yang Lyu(8701121)
output: pdf_document
header-includes: 
  - \renewcommand{\and}{\\}
---

# Introduction
A data set of the 2016 US election polls was given. In this project we aim to understand the data structure by creating various visualizations.  

# Method
We use various R packages to present the data set and to plot the graphs.
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

# Result
```{r}
ElectionPoll<-read.csv("polls_us_election_2016.csv", header = TRUE)
DataDict<-data.frame(
  Variables=colnames(ElectionPoll), Size=nrow(ElectionPoll), 
  Type=sapply(ElectionPoll, function(x) class(x)),
  Number.Unique=sapply(ElectionPoll, function(x) length(unique(x))),
  Number.Missing=sapply(ElectionPoll, function(x) sum(is.na(x))),
  Comment=c("The name of the state where the election is held, “U.S.” means national polls",
            "Start date of poll",
            "End date of poll",
            "Organization name that conducts or analyzes opinion polls",
            "Grade assigned by Fivethirtyeight (sometimes as 538, an American website that focuses on opinion poll) to pollster",
            "Sample size of polls for each pollster",
            "Type of population being polled",
            "Percentage for Hillary Clinton",
            "Percentage for Donald Trump",
            "Percentage for Gary Johnson",
            "Percentage for Evan Mcmullin",
            "Fivethirtyeight adjusted percentage for Hillary Clinton",
            "Fivethirtyeight adjusted percentage for Donald Trump", 
            "Fivethirtyeight adjusted percentage for Gary Johnson",  
            "Fivethirtyeight adjusted percentage for Evan Mcmullin")
)
```

We first create a data variable definition table to give an initial understanding of the data. As we can see, there are a few variables with missing values:
```{r}
DataDict%>%remove_rownames()%>%kable(booktabs = TRUE,caption = "Data Variable Definition")%>%
  kable_styling(font_size=6,latex_options=c("striped","scale_down","hold_position"))%>%
  row_spec(0,bold=TRUE)
```

# Discussion

# Conclusion
