---
title: 'MAT5314 Project 1: Data Visualization'
author:
- Teng Li(7373086)
- Shiya Gao(300381032)
- Chuhan Yue(300376046)
- Yang Lyu(8701121)
output: pdf_document
header-includes: \renewcommand{\and}{\\}
---

# Introduction
A data set of the 2016 US election polls was given. In this project we aim to understand the data structure by creating various visualizations.  

# Method
We use various R packages to present the data set and to plot the graphs.
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(tidyr)
library(tinytex)
```

# Result
We first take a look at the raw data set:
```{r}
ElectionPoll<-read.csv("polls_us_election_2016.csv", header = TRUE)
head(ElectionPoll)
```

As we can see, there are a few variables with missing values:
```{r}
summary(ElectionPoll)
```

## Comparison of candidates' polls in each state

First, we would like to give a brief introduction to the U.S. election system, because it is crucial to understand the background of the data. Voters in each state vote to choose the President of the United States. The candidate who wins the majority of the votes will receive all the electoral votes in that state. Then the sum of the electoral votes in each state is calculated. The total number of electoral votes is 538. The candidate who wins half of the votes plus 1 will win and become the new President of the United States. 

Secondly, we want to analyze the key factors for the candidate’s victory. Since each state has a different number of electoral votes, it is crucial for electors to win in several key states. The reason is that if a candidate wins a certain state, he will win all the electoral votes in that state. So there will be tight competition in states with more votes. 

We want to process the metadata by counting the polls received by each of the four candidates in each state. This result is easier to obtain by multiplying the given size and the proportion. Regardless of the various pollsters, we combine the number of polls for each candidate received in each state although the polls may come from different pollsters. 

```{r, echo=TRUE}
#Create frame contained state, poll number for each candidate
states <- ElectionPoll$state
prop_raw_clinton <- ElectionPoll$rawpoll_clinton
prop_raw_trump <- ElectionPoll$rawpoll_trump
prop_raw_johnson <- ElectionPoll$rawpoll_johnson
prop_raw_mcmullin <- ElectionPoll$rawpoll_mcmullin
size <- ElectionPoll$samplesize

poll_by_state <- data.frame(
  state = states,
  prop_clinton = prop_raw_clinton,
  prop_trump = prop_raw_trump,
  prop_johnson = prop_raw_johnson,
  prop_mcmullin = prop_raw_mcmullin,
  size = size,
  NumVote_clinton = size * (prop_raw_clinton/100),
  NumVote_trump = size * (prop_raw_trump/100),
  NumVote_johnson = size * (prop_raw_johnson/100),
  NumVote_mcmullin = size * (prop_raw_mcmullin/100)
)
poll_by_state[is.na(poll_by_state)] <- 0
head(poll_by_state)
```

```{r, echo=TRUE}
NumVote_State <- cbind(poll_by_state$state, poll_by_state$NumVote_clinton, 
      poll_by_state$NumVote_trump, poll_by_state$NumVote_johnson, 
      poll_by_state$NumVote_mcmullin)
colnames(NumVote_State) <- c("state","Clinton", "Trump", "Johnson", "Mcmullin")
View(NumVote_State)
```

We extracted the variables we were going to use and formed a new data structure  `NumVote_State` with state and four candidates as variables. There are multiple identical values in the State column for a particular state because there are multiple pollsters for each state. In this component, we count the support of each candidate in each state based on the state as the standard, so we ignore the differences in different pollsters in the same state. The combination will take place later.


```{r, echo=TRUE}
NumVoteState <- as.data.frame(NumVote_State) %>%
  pivot_longer(cols = -state,
               names_to = "candidate",
               values_to = "PollNumber")
NumVoteState$PollNumber <- as.numeric(NumVoteState$PollNumber)
View(NumVoteState)
```

We use the `pivot_longer` function to reshape the data and obtain long-format data `NumVoteState`, which is easier to analyze and visualize.

We calculate the total polls received by the four candidates in each state respectively. That is, we combine distinct pollsters if they are in the same state for each candidate. The following shows the support of the four candidates in each state.

- Clinton:

```{r, echo=TRUE}
# Clinton total raw polls by state:
poll_clinton <- filter(NumVoteState, candidate == "Clinton")
Clinton_state <- poll_clinton %>%
  group_by(state) %>%
  summarize(ClintonPolls = sum(PollNumber))
head(Clinton_state)
```

Clinton's total raw polls by state are presented in `Clinton_state`.

- Trump:

```{r, echo=TRUE}
# Trump total raw polls by state:
poll_trump <- filter(NumVoteState, candidate == "Trump")
Trump_state <- poll_trump %>%
  group_by(state) %>%
  summarize(TrumpPolls = sum(PollNumber))
head(Trump_state)
```

Trump's total raw polls by state are presented in `Trump_state `.

- Johnson: 

```{r, echo=TRUE}
# Johnson total raw polls by state:
poll_johnson <- filter(NumVoteState, candidate == "Johnson")
Johnson_state <- poll_johnson %>%
  group_by(state) %>%
  summarize(JohnsonPolls = sum(PollNumber))
head(Johnson_state)
```

Johnson's total raw polls by state are presented in `Johnson_state`.

- Mcmullin: 

```{r, echo=TRUE}
# Mcmullin total raw polls by state:
poll_mcmullin <- filter(NumVoteState, candidate == "Mcmullin")
Mcmullin_state <- poll_mcmullin %>%
  group_by(state) %>%
  summarize(McmullinPolls = sum(PollNumber))
head(Mcmullin_state)
```

Mcmullin's total raw polls by state are presented in `Mcmullin_state`.

## Visualization of the poll proportion of the four candidates in each state:

```{r g_11, echo=TRUE, fig.cap = "Poll Percentage by States."}
g_11 <- ggplot(data = NumVoteState, mapping = aes( x = state, fill = candidate)) +
  geom_col(aes(y = PollNumber),position='fill') +
  labs(x = "State", y = "Percentage of Polls", 
       title = "Poll Percentage by States", 
       caption = "Data: NumVoteState") +
  scale_fill_manual(values=c("Clinton" = "blue", "Trump" = "red", 
                             "Johnson" = "green", "Mcmullin" = "yellow")) +
  theme(axis.text.x = element_text(angle = 50, size = 6, vjust = 0.5)) + 
  theme(axis.title.x = element_text(size = 9, vjust = 0),
        axis.title.y = element_text(size = 9, vjust = 3)) + 
  theme(plot.title = element_text(size = 12, face = "bold", 
                                  margin = margin(0, 0, 0, 0))) +
  theme(legend.position = "bottom")
print(g_11)
```

Figure \@ref(fig:g_11) shows the poll proportions of the four candidates in each state distinguished by colors. We can clearly observe which candidate is likely to win all the electoral votes in each state, which is helpful in estimating the outcome of the presidential election.

---

## Poll Distribution

We are eager to observe the distribution of polls among the four candidates. 

# Discussion

# Conclusion
